---
title: "Final Project"
author: "Tang Ching Xian"
date: "2023-11-21"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.


```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(highcharter)
library(gt)
library(htmltools)
library(viridis)
library(maps)
library(sf)
```

```{r}
## Libraries 
library(tidyverse)
library(maps)
library(sf)


createInteractiveMap <- function(data_file, map_resolution = "world") {
  library(ggplot2)
  library(dplyr)
  library(plotly)

  # Import the data with coordinates
  world_map <- map_data(map_resolution)
  accesstonet <- read.csv(data_file)
  map_data <- accesstonet %>% rename(region = country)

  map <- left_join(world_map, map_data, by = "region")

  # Create the ggplot object
  gg <- ggplot(map, aes(long, lat, group = group, text = paste("Country: ", region, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal()

  # Convert ggplot to plotly
  map_plotly <- ggplotly(gg, tooltip = "text")

  # Display the interactive map
  return(map_plotly)
}

# Usage
interactiveMap <- createInteractiveMap("food-waste-by-country-2023.csv")
interactiveMap


```
=================================================================================


-----------------------------------------------------------------------

### Dataset

```{r}
# Libraries
library (tidyverse)
library (plotly)
library (shiny)
library(dplyr)

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

energy_stat <- read.csv('waste_energy_stat.csv')


# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Filtering waste_03_22 dataset for specific years and waste types
wasteselected <- clean_waste_03_22 %>%
  filter(year %in% 2003:2022,
         waste_type %in% c("Plastics", "Ferrous Metals", "Non-Ferrous Metals", "Glass", "Food"))%>%
  arrange(waste_type)

  
### Calculating and adding recycling rate into DataFrame for analysis.

# Calculate recycling_rate and round to 2 decimal places
clean_waste_03_22$recycling_rate <- round(
  clean_waste_03_22$total_waste_recycled_tonne / clean_waste_03_22$total_waste_generated_tonne,2
)

### Calculating and adding wasting rate into DataFrame for analysis.

# Calculate wasting_rate and round to 2 decimal places
clean_waste_03_22$wasting_rate <- round(
  1-clean_waste_03_22$recycling_rate,2
)

# Display the cleaned data
clean_waste_03_22



```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}
# Colors
custom_colors <- viridis::mako(n=30)

# Grouping by material and summarizing total waste and recycled
summary_df <- clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated))

# Adding relevant headings
colnames(summary_df) <- c("Material", "Total Waste Generated (tonnes)", "Total Waste Recycled (tonnes)")

# Displaying the summarized data
summary_df

# Adapted code for waste data
clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated)) %>%
  head(15) 

```

### Chart C

```{r}
# Bar Plot for Waste Generated and Waste Recycled
fig <- plot_ly(
  x = clean_waste_03_22$year,
  y = clean_waste_03_22$total_waste_generated_tonne,
  type = "bar",
  name = "Waste Generated"
) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_recycled_tonne,
    type = "bar",
    name = "Waste Recycled"
  ) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_not_recycled_tonne,
    type = "bar",
    name = "Waste Not Recycled"
  )

fig
```

### Chart D

```{r}
#use merge function: 
total_data <- merge(clean_waste_03_22, energy_stat, by.x = "waste_type", by.y = "Material", all.x = TRUE)
total_data$energy_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$energy_saved)
total_data$crude_oil_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$crude_oil_saved)

# Display the first few rows
head(total_data)

# Box plot of total_waste_not_recycled_tonne
fig <- plot_ly(total_data, x = ~year, y = ~total_waste_recycled_tonne, type = "box")
fig <- fig %>% layout(title = "Box Plot of Total Waste Not Recycled Tonne", yaxis = list(title = "Total Waste Not Recycled Tonne"))

fig

# Box plot of total_energy_saved
fig2 <- plot_ly(total_data, x = ~year, y = ~energy_saved, type = "box")
fig2 <- fig %>% layout(title = "Box Plot of Total Energy Saved", yaxis = list(title = "Total Energy Saved"))

fig2
```
=======================================================================
### Chart E

#```{r}
#scatter plot graph
fig <- plot_ly(
  data = total_data,
  x = ~year,
  y = ~energy_saved,
  size = ~total_waste_generate_tonne,
  color = ~material,
  sizes = c(10, 60),
  type = "scatter",
  mode = "markers"
)

fig


# Energy saved per year
annual_energy_savings <- total_data %>%
  group_by(year) %>%
  summarize(energy_saved = sum(energy_saved))

# Convert total_energy_saved to GWh
annual_energy_savings$total_energy_saved <- round(annual_energy_savings$energy_saved / 1000000, 2)

# Display the result
tail(annual_energy_savings)
```
