---
title: "No Time to Waste: Food loss and waste must be reduced for greater food security and environmental sustainability. Turn Food Waste into Resourceful waste!"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---
```{r, results='asis', echo=FALSE}
cat("
<section class=\"image-container\">
  <img src=\"src/delicious-vietnamese-food-including-pho-ga-noodles-spring-rolls-white-table.jpg\" style=\"width:100%;height:300px;object-fit:cover;\"/>
  <h5 class=\"center-text-bg-img\">Unveiling the true severity of food waste</h5>
  <p class=\"bottom-right-text-bg-img\">Image by [Freepik](https://www.freepik.com/free-photo/close-up-hand-holding-wooden-board_14351726.htm#query=food%20waste&position=26&from_view=search&track=ais)</p>
</section>

<section id=\"introduction\" class=\"container\" width=\"100%\">
  <h4 class=\"text-center fw-bold my-4\">Introduction</h4>
  <p>Food comes from a long journey, starting with the farmer and ending up in our homes. This journey, called the supply chain, uses a lot of the Earth's resources. When we sit down to eat, we get to enjoy the amazing result. Tradition and good cooking bring us tasty, nourishing, joyful, passionate, and life-giving meals. However, the sad truth is that we throw away a ton of our food—loads of it! And when we do that, we're also tossing away all the valuable resources that were essential to make it in the first place.</p>

  <p>**WASTING FOOD IS WASTING AGRICULTURAL LAND.** The food produced but wasted globally every year requires 1.4 billion hectares of land, which is close to 30% of the world agricultural land area (UNEP, 2022).</p>

  <p>**WASTING FOOD IS CONTRIBUTING TO GLOBAL WARMING.** The carbon footprint of food wasted globally every year is 3.3 billion tonnes of carbon dioxide. If food waste was a country, it would be the third top emitter after the USA and China (FAO, 2011).</p>

  <p>**WASTING FOOD IS WASTING MONEY.** Besides its environmental cost, food wastage also represents a loss of economic value. On a global scale, the economic cost of the overall amount of food wastage is about $US 750 billion, equivalent to the GDP of Switzerland (FAO, 2011).</p>

  <h4 id=\"Wondering how much food we waste\" class=\"text-center fw-bold my-4\">How much **food is wasted?**</h4>
  <p>According to the **United Nations' Food and Agricultural Organisation of United Nation (FAO)**, an estimation of all the food produced every year for human consumption in the world, one-third does not reach our stomachs.</p>
</section>
")
```


```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(highcharter)
library(gt)
library(htmltools)
library(viridis)
library(maps)
library(sf)
library(shiny)
```

```{=html}
<p style="text-align: center; font-weight: bold;">Let's explore the global landscape on a world map to understand the varying levels of food waste generated in each country.</p>
```

```{r}


```

```{r}
# Libraries
library(tidyverse)
library(maps)
library(sf)
library(shiny)
library(plotly)

# Function to create the interactive map
createInteractiveMap <- function(selected_country) {
  # Import the data with coordinates
  world_map <- map_data("world")
  accesstonet <- read.csv("food-waste-by-country-2023.csv")
  map_data <- accesstonet %>% rename(country = Country)

  map <- left_join(world_map, map_data, by = "country")

  # Create the ggplot object
  gg <- ggplot(map, aes(x = long, y = lat, group = group, text = paste("Country: ", country, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal()

  # Convert ggplot to plotly
  map_plotly <- ggplotly(gg, tooltip = "text")

  # Set initial zoom to the selected country
  if (!is.null(selected_country)) {
    map_plotly$x$layout$geo$fitbounds <- "locations"
    map_plotly$x$data[[1]]$locations <- selected_country
  }

  # Display the interactive map
  return(map_plotly)
}

# Shiny app
ui <- fluidPage(
  selectInput("selected_country", "Select Country", choices = ""),
  plotlyOutput("interactive_map")
)

server <- function(input, output, session) {
  observe({
    choices <- unique(read.csv("food-waste-by-country-2023.csv")$Country)
    updateSelectInput(session, "selected_country", choices = choices)
  })

  # Update the map based on the selected country
  output$interactive_map <- renderPlotly({
    selected_country_data <- filter(read.csv("food-waste-by-country-2023.csv"), country == input$selected_country)
    map_plot <- createInteractiveMap(selected_country_data$country)
    map_plot
  })
}

shinyApp(ui, server)
```

```{r}
# Libraries
library(tidyverse)
library(maps)
library(sf)
library(shiny)
library(plotly)

# Function to create the interactive map
createInteractiveMap <- function(selected_country) {
  # Import the data with coordinates
  world_map <- map_data("world")
  accesstonet <- read.csv("food-waste-by-country-2023.csv")
  map_data <- accesstonet %>% rename(country = Country)

  map <- left_join(world_map, map_data, by = "country")

  # Create the ggplot object
  gg <- ggplot(map, aes(x = long, y = lat, group = group, text = paste("Country: ", country, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal() +
    theme(legend.position = "top")  # Set legend position to top

  # Convert ggplot to plotly
  map_plotly <- ggplotly(gg, tooltip = "text")

  # Set initial zoom to the selected country
  if (!is.null(selected_country)) {
    map_plotly$x$layout$geo$fitbounds <- "locations"
    map_plotly$x$data[[1]]$locations <- selected_country
  }

  # Display the interactive map
  return(map_plotly)
}

# Shiny app
ui <- fluidPage(
  selectInput("selected_country", "Select Country", choices = ""),
  plotlyOutput("interactive_map")
)

server <- function(input, output, session) {
  observe({
    choices <- unique(read.csv("food-waste-by-country-2023.csv")$Country)
    updateSelectInput(session, "selected_country", choices = choices)
  })

  # Update the map based on the selected country
  output$interactive_map <- renderPlotly({
    selected_country_data <- filter(read.csv("food-waste-by-country-2023.csv"), country == input$selected_country)
    map_plot <- createInteractiveMap(selected_country_data$country)
    map_plot
  })
}

shinyApp(ui, server)
```

```{r}

library(shiny)
library(shinydashboard)
library(plotly)
library(tidyverse)

# Load your data
fwd_data <- read.csv("FWD.csv")

# UI
ui <- dashboardPage(
  dashboardHeader(title = "Interactive Boxplot and Heatmap"),
  dashboardSidebar(),
  dashboardBody(
    box(title = "Boxplot by Stage", 
        plotlyOutput("boxplot")),
    
    box(title = "Heatmap of Available Data", 
        plotlyOutput("heatmap"))
  )
)

# Server
server <- function(input, output) {
  
  # Reactive subset of data for selected stages
  selected_stages <- reactive({
    fwd_data %>%
      select("Stage", "Retail", "FoodServices", "Households")
  })
  
  # Boxplot
  output$boxplot <- renderPlotly({
    plot_ly(selected_stages(), 
            y = ~Retail, 
            type = "box", 
            name = "Retail") %>%
      add_trace(y = ~FoodServices, name = "Food Services") %>%
      add_trace(y = ~Households, name = "Households") %>%
      layout(title = "Boxplot by Stage",
             xaxis = list(title = "Stage"),
             yaxis = list(title = "Value"))
  })
  
  # Heatmap
  output$heatmap <- renderPlotly({
    plot_ly(z = ~selected_stages()$Retail,
            x = ~selected_stages()$Stage,
            y = ~c("Retail")) %>%
      add_trace(z = ~selected_stages()$FoodServices, y = ~c("Food Services")) %>%
      add_trace(z = ~selected_stages()$Households, y = ~c("Households")) %>%
      layout(title = "Heatmap of Available Data",
             xaxis = list(title = "Stage"),
             yaxis = list(title = ""),
             barmode = "stack")
  })
}

# Run the Shiny app
shinyApp(ui, server)
```

###Top 5 countries and Bottom 5 countries with combined per capita food waste
```{r setup, include=FALSE}
library(ggplot2)
library(stringr)
library(grid)
library(ggchicklet)

FWD <- read.csv("FWD.csv")

# Order the data by combined_estimate_kpcpy in descending order
topCountries <- FWD[order(-FWD$combined_estimate_kpcpy), ]

# Select the top 5 countries
top5Countries <- topCountries[1:5, ]

ggplot(top5Countries, aes(x = reorder(country, combined_estimate_kpcpy), y = combined_estimate_kpcpy)) +
  geom_chicklet(radius = grid::unit(2, 'mm'), fill = "#d62023") +
  geom_text(aes(label = combined_estimate_kpcpy), hjust = 1.8, vjust = 0.5, size = 3, colour = "white", fontface = "bold") +  
  labs(title = str_wrap("The 5 Countries with Highest Combined Per-Capita Food Waste", width = 40),
       subtitle = "Per-Capita Food Waste (kg/year)",
       x = "", y = "Per-Capita Food Waste (kg/year)") +
  remove_background_theme() +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold", margin = margin(b = 20, l = 20, r = 20)),
        plot.subtitle = element_text(hjust = 0.5, size = 14)) 

# Select the bottom 5 countries
bottom5Countries <- topCountries[(nrow(topCountries)-4):nrow(topCountries), ]

# Show the 5 countries with the lowest combined per-capita food waste
ggplot(bottom5Countries, aes(x = reorder(country, combined_estimate_kpcpy), y = combined_estimate_kpcpy)) +
  geom_chicklet(radius = grid::unit(2, 'mm'), fill = "#3fb619") +
  geom_text(aes(label = combined_estimate_kpcpy), hjust = 1.8, vjust = 0.5, size = 3, colour = "white", fontface = "bold") +
  labs(title = str_wrap("The 5 Countries with Lowest Combined Per-Capita Food Waste", width = 40),
       x = "", y = "Per-Capita Food Waste (kg/year)") +
  remove_background_theme() +
  coord_flip() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold", margin = margin(b = 20, l = 20, r = 20)))

```

```{r}
# Load necessary libraries
library(readr)
library(ggplot2)

# Read the CSV file
FWD_UNEP <- read_csv("FWD.csv")

# Function to calculate total estimates for each sector
calculate_total <- function(data, column) {
  return(sum(data[[column]]))
}

# Calculate total estimates for each sector
total_household <- calculate_total(FWD_UNEP, "household_estimate_tpy")
total_retail <- calculate_total(FWD_UNEP, "retail_estimate_tpy")
total_food_service <- calculate_total(FWD_UNEP, "food_service_estimate_tpy")

# Create a data frame with total estimates
total_FW_by_sector <- data.frame(
  category = c("Household", "Retail", "Food Service"),
  value = c(total_household, total_retail, total_food_service),
  color = c("#d62023", "#3fb619", "#8FB8FDff")
)

# Function to create bar and pie charts
create_chart <- function(data, x, y, fill, title, x_label, y_label, coord_type = NULL) {
  ggplot(data, aes_string(x = x, y = y, fill = fill)) +
    geom_bar(stat = "identity", color = "black", width = 0.7) +
    labs(title = title, x = x_label, y = y_label) +
    scale_fill_manual(values = c("#d62023", "#3fb619", "#8FB8FDff")) +
    if (!is.null(coord_type)) coord_type else NULL
}

# Create bar chart
bar_chart <- create_chart(total_FW_by_sector, "category", "value", "category",
                           "Total Food Waste by Sector", "Sector", "Total Food Waste (tpy)", NULL)

# Create pie chart
pie_chart <- create_chart(total_FW_by_sector, "", "value", "category",
                           "Total Food Waste Distribution", "Sector", "Sector",
                           coord_type = coord_polar("y"))

# Display the charts
print(bar_chart)
print(pie_chart)
```
The chart shows that households are the largest contributors to food waste, accounting for 61.1% of the total. This is followed by food services (26.2%) and retail establishments (12.7%).

#intend to able to zoom into country when choosing the country,default singapore

### Dataset

```{r}

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

energy_stat <- read.csv('waste_energy_stat.csv')


# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Filtering waste_03_22 dataset for specific years and waste types
wasteselected <- clean_waste_03_22 %>%
  filter(year %in% 2003:2022,
         waste_type %in% c("Plastics", "Ferrous Metals", "Non-Ferrous Metals", "Glass", "Food"))%>%
  arrange(waste_type)

  
### Calculating and adding recycling rate into DataFrame for analysis.

# Calculate recycling_rate and round to 2 decimal places
clean_waste_03_22$recycling_rate <- round(
  clean_waste_03_22$total_waste_recycled_tonne / clean_waste_03_22$total_waste_generated_tonne,2
)

### Calculating and adding wasting rate into DataFrame for analysis.

# Calculate wasting_rate and round to 2 decimal places
clean_waste_03_22$wasting_rate <- round(
  1-clean_waste_03_22$recycling_rate,2
)

# Display the cleaned data
head(clean_waste_03_22)
```


```{r}
# Libraries

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Define UI for the Shiny app
ui <- fluidPage(
  titlePanel("Waste Visualization App"),
  sidebarLayout(
    sidebarPanel(
      selectInput("material", "Choose Material:", 
                  choices = unique(clean_waste_03_22$waste_type),
                  selected = "Plastics")
    ),
    mainPanel(
      plotlyOutput("waste_plot")
    )
  )
)

# Define server logic for the Shiny app
server <- function(input, output) {
  
  # Filter data based on selected material
  selected_data <- reactive({
    clean_waste_03_22 %>%
      filter(waste_type == input$material)
  })
  
  # Render the plot
  output$waste_plot <- renderPlotly({
    gg <- ggplot(selected_data(), aes(x = year)) +
      geom_line(aes(y = total_waste_generated_tonne, color = "Waste Generated"), size = 1) +
      geom_line(aes(y = total_waste_recycled_tonne, color = "Waste Recycled"), size = 1) +
      geom_line(aes(y = total_waste_not_recycled_tonne, color = "Waste Not Recycled"), size = 1) +
      labs(title = paste("Waste Over Years -", input$material),
           x = "Year", y = "Tonnes") +
      theme_minimal()
    
    ggplotly(gg, tooltip = "text")
  })
}

# Run the Shiny app
shinyApp(ui, server)
```
```{r}

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Filtering waste_03_22 dataset for specific years and waste types
wasteselected <- clean_waste_03_22 %>%
  filter(year %in% 2003:2022,
         waste_type %in% c("Plastics", "Ferrous Metals", "Non-Ferrous Metals", "Glass", "Food")) %>%
  arrange(waste_type)

  
### Calculating and adding recycling rate into DataFrame for analysis.

# Calculate recycling_rate and round to 2 decimal places
clean_waste_03_22$recycling_rate <- round(
  clean_waste_03_22$total_waste_recycled_tonne / clean_waste_03_22$total_waste_generated_tonne, 2
)

### Calculating and adding wasting rate into DataFrame for analysis.

# Calculate wasting_rate and round to 2 decimal places
clean_waste_03_22$wasting_rate <- round(
  1 - clean_waste_03_22$recycling_rate, 2
)

# Define UI for the Shiny app
ui <- fluidPage(
  titlePanel("Recycling and Wasting Rates Visualization App"),
  sidebarLayout(
    sidebarPanel(
      selectInput("material", "Choose Material:",
                  choices = unique(clean_waste_03_22$waste_type),
                  selected = "Plastics")
    ),
    mainPanel(
      plotlyOutput("rate_plot")
    )
  )
)

# Define server logic for the Shiny app
server <- function(input, output) {
  
  # Filter data based on selected material
  selected_data <- reactive({
    clean_waste_03_22 %>%
      filter(waste_type == input$material)
  })
  
  # Render the plot
  output$rate_plot <- renderPlotly({
    gg <- ggplot(selected_data(), aes(x = year)) +
      geom_line(aes(y = recycling_rate, color = "Recycling Rate"), size = 1) +
      geom_line(aes(y = wasting_rate, color = "Wasting Rate"), size = 1) +
      labs(title = paste("Recycling and Wasting Rates Over Years -", input$material),
           x = "Year", y = "Rate") +
      theme_minimal()
    
    ggplotly(gg, tooltip = "text")
  })
}

```


### Chart B

```{r}
# Colors
custom_colors <- viridis::mako(n=30)

# Grouping by material and summarizing total waste and recycled
summary_df <- clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated))

# Adding relevant headings
colnames(summary_df) <- c("Material", "Total Waste Generated (tonnes)", "Total Waste Recycled (tonnes)")

# Displaying the summarized data
summary_df

# Adapted code for waste data
clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated)) %>%
  head(15) 

clean_waste_03_22 <- clean_waste_03_22 %>% 
  pivot_longer(cols = disposed_tonnes:generated_tonnes,
              names_to = "Sector",
              values_to = "Tonnes") %>% 
  mutate(Sector = case_when(
    Sector == "disposed_tonnes" ~ "Disposed",
    Sector == "generated_tonnes" ~ "Generated",
    Sector == "recycled_tonnes" ~ "Recycled",
    TRUE ~ as.character(Sector)
  ))

# Rename observation rows in the "Sector" column

# Spline Plot
clean_waste_03_22 %>% 
  hchart('spline', hcaes(x = 'year', y = 'Tonnes', group = "Sector")) %>% 
  hc_title(text = "Waste Statistics and Overall Recycling Table in Singapore")


```

### Chart C

```{r}
# Bar Plot for Waste Generated and Waste Recycled
fig <- plot_ly(
  x = clean_waste_03_22$year,
  y = clean_waste_03_22$total_waste_generated_tonne,
  type = "bar",
  name = "Waste Generated"
) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_recycled_tonne,
    type = "bar",
    name = "Waste Recycled"
  ) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_not_recycled_tonne,
    type = "bar",
    name = "Waste Not Recycled"
  )

fig
```

### Chart D

```{r}
#use merge function: 
total_data <- merge(clean_waste_03_22, energy_stat, by.x = "waste_type", by.y = "Material", all.x = TRUE)
total_data$energy_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$energy_saved)
total_data$crude_oil_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$crude_oil_saved)

# Display the first few rows
head(total_data)

# Box plot of total_waste_not_recycled_tonne
fig <- plot_ly(total_data, x = ~year, y = ~total_waste_recycled_tonne, type = "box")
fig <- fig %>% layout(title = "Box Plot of Total Waste Not Recycled Tonne", yaxis = list(title = "Total Waste Not Recycled Tonne"))

fig

# Box plot of total_energy_saved
fig2 <- plot_ly(total_data, x = ~year, y = ~energy_saved, type = "box")
fig2 <- fig2 %>% layout(title = "Box Plot of Total Energy Saved", yaxis = list(title = "Total Energy Saved"))

fig2
```

### Chart E

```{r}
first_5_rows <- head(clean_waste_03_22, 5) %>% select(waste_type, generated_tonnes)

grouped_sum <- clean_waste_03_22 %>%
  slice(6:14) %>%
  summarise(waste_type = "Others", generated_tonnes = sum(generated_tonnes))

final_result <- bind_rows(first_5_rows, grouped_sum)

highchart() %>%
  hc_chart(type = "pie") %>%
  hc_add_series(final_result, "pie", hcaes(y = "generated_tonnes", name = "waste_type", color = "waste_type"), center = c(50, 50), 
                innerSize="50%") %>% 
  hc_tooltip(
    pointFormat = "<b>{point.name}</b>: {point.y} tonnes/year ({point.percentage:.1f}%)"
  ) %>%  hc_plotOptions(
    pie = list(
      dataLabels = list(
        enabled = TRUE,
        format = "<b>{point.name}</b>:<br>({point.percentage:.1f}%)",
        style = list(fontSize = "10px", fontWeight = "bold")
      )
    ),
    showInLegend = TRUE
  ) %>% 
  hc_title(text = "Total Generated Waste by Type in Singapore") %>%
  hc_legend(enabled = TRUE) %>%
  hc_subtitle(text = "Total Food Waste (Tonnes/Year)")

```

```{r}
# Assuming 'recycling_rate' is the column in clean_waste_03_22 containing recycling rates
first_5_rows <- head(clean_waste_03_22, 5) %>% select(waste_type, generated_tonnes, recycling_rate)

grouped_sum <- clean_waste_03_22 %>%
  slice(6:14) %>%
  summarise(waste_type = "Others", 
            generated_tonnes = sum(generated_tonnes),
            recycling_rate = mean(recycling_rate))

final_result <- bind_rows(first_5_rows, grouped_sum)

highchart() %>%
  hc_chart(type = "pie") %>%
  hc_add_series(final_result, "pie", 
                hcaes(y = "generated_tonnes", name = "waste_type", color = "waste_type"), 
                center = c(50, 50), 
                innerSize="50%") %>% 
  hc_tooltip(
    pointFormat = "<b>{point.name}</b>: {point.y} tonnes/year ({point.percentage:.1f}%)<br>Recycling Rate: {point.recycling_rate}%"
  ) %>%  
  hc_plotOptions(
    pie = list(
      dataLabels = list(
        enabled = TRUE,
        format = "<b>{point.name}</b>:<br>({point.percentage:.1f}%)",
        style = list(fontSize = "10px", fontWeight = "bold")
      )
    ),
    showInLegend = TRUE
  ) %>% 
  hc_title(text = "Total Generated Waste by Type in Singapore") %>%
  hc_legend(enabled = TRUE) %>%
  hc_subtitle(text = "Total Food Waste (Tonnes/Year) and Recycling Rate")
```


As per information from the National Environment Agency (NEA) (n.d.), Singapore produced a staggering 763,000 tonnes of food waste in 2022, constituting 11% of the total waste generated. This quantity has exhibited a consistent rise over the years, and the trend is anticipated to persist with the ongoing expansion of the population and economy. Moreover, our discussion has revealed that the recycling rate for food waste is notably low, further emphasizing the significant contribution of food waste to the overall waste scenario in Singapore.

### Chart F



```{r}
#scatter plot graph
#use merge function: 
total_data <- merge(clean_waste_03_22, energy_stat, by.x = "waste_type", by.y = "Material", all.x = TRUE)
total_data$energy_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$energy_saved)
total_data$crude_oil_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$crude_oil_saved)

fig3 <- plot_ly(
  data = total_data,
  x = ~year,
  y = ~energy_saved,
  size = ~total_waste_generate_tonne,
  color = ~material,
  sizes = c(10, 60),
  type = "scatter",
  mode = "markers"
)

fig3


# Energy saved per year
annual_energy_savings <- total_data %>%
  group_by(year) %>%
  summarize(energy_saved = sum(energy_saved))

# Convert total_energy_saved to GWh
annual_energy_savings$total_energy_saved <- round(annual_energy_savings$energy_saved / 1000000, 2)

# Display the result
tail(annual_energy_savings)
```

```{r}
# References
cat("
<section id=\"references\" class=\"container mt-5\" style=\"width:100%;\">

<div class=\"accordion\" id=\"accordionExample\">
  <div class=\"accordion-item\">
    <h2 class=\"accordion-header\" id=\"headingOne\">
      <button class=\"accordion-button\" type=\"button\" data-bs-toggle=\"collapse\" data-bs-target=\"#collapseOne\" aria-expanded=\"f\" aria-controls=\"collapseOne\">
        <h4>References</h4>
      </button>
    </h2>
    <div id=\"collapseOne\" class=\"accordion-collapse collapse show\" aria-labelledby=\"headingOne\" data-bs-parent=\"#accordionExample\">
      <div class=\"accordion-body\">
        <p>Miquelis, P. (2019, january 25). The Truth About Food Waste in Singapore (and beyond). <a href=\"https://greenisthenewblack.com/the-truth-about-food-waste-in-singapore-and-beyond/\" target=\"_blank\">https://greenisthenewblack.com/the-truth-about-food-waste-in-singapore-and-beyond/</a></p> 
        
<p>United Nation Environment Programme. (2022, September 26). Why the global fight to tackle food waste has only just begun. <a href=\"https://www.unep.org/news-and-stories/story/why-global-fight-tackle-food-waste-has-only-just-begun\" target=\"_blank\">https://www.unep.org/news-and-stories/story/why-global-fight-tackle-food-waste-has-only-just-begun</a></p>

<p>https://worldpopulationreview.com/country-rankings/food-waste-by-country</p>

<p>Lee, H.W. (2022, January 3).Food Wastage Statistics in Singapore. <a href=\"https://www.bestinsingapore.co/food-wastage-statistics-singapore/\" target=\"_blank\">https://www.bestinsingapore.co/food-wastage-statistics-singapore/</a></p>

<p>Food and Agriculture Organization. (n.d.). The State of Food and Agriculture - 2019. <a href=\"https://www.fao.org/state-of-food-agriculture/2019/en/\" target=\"_blank\">https://www.fao.org/state-of-food-agriculture/2019/en/</a></p>

<p>Low, Y.J. (2019, August 6). Every Singaporean household throws away an average S$258 worth of food a year: Study. <a href=\"https://www.todayonline.com/singapore/every-singaporean-household-throws-away-average-s258-worth-food-year-study\" target=\"_blank\">https://www.todayonline.com/singapore/every-singaporean-household-throws-away-average-s258-worth-food-year</a></p>

<p>Food and Agriculture Organization of the United Nations (FAO). 2011. Global food losses and food waste – Extent, causes and prevention. Rome. <a href=\"https://www.fao.org/3/mb060e/mb060e00.htm\" target=\"_blank\">https://www.fao.org/3/mb060e/mb060e00.htm</a></p>

<p>National Environment Agency (NEA). (n.d.). Waste Statistics and Overall Recycling. <a href=\"https://www.nea.gov.sg/our-services/waste-management/waste-statistics-and-overall-recycling\" target=\"_blank\">https://www.nea.gov.sg/our-services/waste-management/waste-statistics-and-overall-recycling</a></p>

<p>Towards Zero Waste. (n.d.). Food Waste. <a href=\"https://www.towardszerowaste.gov.sg/foodwaste/\" target=\"_blank\">https://www.towardszerowaste.gov.sg/foodwaste/</a></p>

<p>O'Neill, A. (2023, February 2). Gross domestic product (GDP) per capita in selected global regions at current prices in 2021. <a href=\"https://www.statista.com/statistics/256413/gross-domestic-product-per-capita-in-selected-global-regions/\" target=\"_blank\">https://www.statista.com/statistics/256413/gross-domestic-product-per-capita-in-selected-global-regions/</a></p>
      </div>
    </div>
  </div>
</div>
</section> <!-- End of References Section -->
</section>
</div>
</section>
</section>
</section>
")
```


