---
title: "No Time to Waste: Food loss and waste must be reduced for greater food security and environmental sustainability. Turn Food Waste into Resourceful waste!"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---
```{r, results='asis', echo=FALSE}
cat("
<section class=\"image-container\">
  <img src=\"src/delicious-vietnamese-food-including-pho-ga-noodles-spring-rolls-white-table.jpg\" style=\"width:100%;height:300px;object-fit:cover;\"/>
  <h5 class=\"center-text-bg-img\">Unveiling the true severity of food waste</h5>
  <p class=\"bottom-right-text-bg-img\">Image by [Freepik](https://www.freepik.com/free-photo/close-up-hand-holding-wooden-board_14351726.htm#query=food%20waste&position=26&from_view=search&track=ais)</p>
</section>

<section id=\"introduction\" class=\"container\" width=\"100%\">
  <h4 class=\"text-center fw-bold my-4\">Introduction</h4>
  <p>Food comes from a long journey, starting with the farmer and ending up in our homes. This journey, called the supply chain, uses a lot of the Earth's resources. When we sit down to eat, we get to enjoy the amazing result. Tradition and good cooking bring us tasty, nourishing, joyful, passionate, and life-giving meals. However, the sad truth is that we throw away a ton of our food—loads of it! And when we do that, we're also tossing away all the valuable resources that were essential to make it in the first place.</p>

  <p>**WASTING FOOD IS WASTING AGRICULTURAL LAND.** The food produced but wasted globally every year requires 1.4 billion hectares of land, which is close to 30% of the world agricultural land area (UNEP, 2022).</p>

  <p>**WASTING FOOD IS CONTRIBUTING TO GLOBAL WARMING.** The carbon footprint of food wasted globally every year is 3.3 billion tonnes of carbon dioxide. If food waste was a country, it would be the third top emitter after the USA and China (FAO, 2011).</p>

  <p>**WASTING FOOD IS WASTING MONEY.** Besides its environmental cost, food wastage also represents a loss of economic value. On a global scale, the economic cost of the overall amount of food wastage is about $US 750 billion, equivalent to the GDP of Switzerland (FAO, 2011).</p>

  <h4 id=\"Wondering how much food we waste\" class=\"text-center fw-bold my-4\">How much **food is wasted?**</h4>
  <p>According to the **United Nations' Food and Agricultural Organisation of United Nation (FAO)**, an estimation of all the food produced every year for human consumption in the world, one-third does not reach our stomachs.</p>
</section>
")
```


```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(highcharter)
library(gt)
library(htmltools)
library(viridis)
library(maps)
library(sf)
library(shiny)
```

```{r}
## Libraries 
library(tidyverse)
library(maps)
library(sf)
library(shiny)


createInteractiveMap <- function(data_file, map_resolution = "world") {
  library(ggplot2)
  library(dplyr)
  library(plotly)

  # Import the data with coordinates
  world_map <- map_data(map_resolution)
  accesstonet <- read.csv(data_file)
  map_data <- accesstonet %>% rename(region = country)

  map <- left_join(world_map, map_data, by = "Country")

  # Create the ggplot object
  gg <- ggplot(map, aes(long, lat, group = group, text = paste("Country: ", country, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal()

  # Convert ggplot to plotly
  map_plotly <- ggplotly(gg, tooltip = "text")

  # Display the interactive map
  return(map_plotly)
}

## Libraries 
library(tidyverse)
library(maps)
library(sf)
library(plotly)

createInteractiveMap <- function(data_file, map_resolution = "world") {
  # Import the data with coordinates
  world_map <- map_data(map_resolution)
  accesstonet <- read.csv(data_file)
  map_data <- accesstonet %>% rename(region = country)

  map <- left_join(world_map, map_data, by = "region")

  # Create the ggplot object
  gg <- ggplot(map, aes(long, lat, group = group, text = paste("Country: ", region, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal()

  # Convert ggplot to plotly with specified width and height
  map_plotly <- ggplotly(gg, tooltip = "text", width = 1000, height = 600)

  # Set scrollZoom to FALSE
  map_plotly$x$layout$updatemenus <- list(list(type = "zoom",
showactive = FALSE,
buttons = list(list(args = list("zoomType", "zoomin"),
label = "Zoom In",
method = "relayout"),
list(args = list("zoomType", "zoomout"),
label = "Zoom Out",
method = "relayout")),
direction = "left",
pad = list(r = 10, t = 87),
showactive = FALSE,
type = "buttons",
x = 0.1,
y = 1))

  # Display the interactive map
  return(map_plotly)
}

# Usage
interactiveMap <- createInteractiveMap("food-waste-by-country-2023.csv")
interactiveMap
```


# Load necessary libraries
library(readr)
library(ggplot2)

# Read the CSV file
FWD_UNEP <- read_csv("FWD.csv")

# Function to calculate total estimates for each sector
calculate_total <- function(data, column) {
  return(sum(data[[column]]))
}

# Calculate total estimates for each sector
total_household <- calculate_total(FWD_UNEP, "household_estimate_tpy")
total_retail <- calculate_total(FWD_UNEP, "retail_estimate_tpy")
total_food_service <- calculate_total(FWD_UNEP, "food_service_estimate_tpy")

# Create a data frame with total estimates
total_FW_by_sector <- data.frame(
  category = c("Household", "Retail", "Food Service"),
  value = c(total_household, total_retail, total_food_service),
  color = c("#d62023", "#3fb619", "#8FB8FDff")
)

# Function to create bar and pie charts
create_chart <- function(data, x, y, fill, title, x_label, y_label, coord_type = NULL) {
  ggplot(data, aes_string(x = x, y = y, fill = fill)) +
    geom_bar(stat = "identity", color = "black", width = 0.7) +
    labs(title = title, x = x_label, y = y_label) +
    scale_fill_manual(values = c("#d62023", "#3fb619", "#8FB8FDff")) +
    if (!is.null(coord_type)) coord_type else NULL
}

# Create bar chart
bar_chart <- create_chart(total_FW_by_sector, "category", "value", "category",
                           "Total Food Waste by Sector", "Sector", "Total Food Waste (tpy)", NULL)

# Create pie chart
pie_chart <- create_chart(total_FW_by_sector, "", "value", "category",
                           "Total Food Waste Distribution", "Sector", "Sector",
                           coord_type = coord_polar("y"))

# Display the charts
print(bar_chart)
print(pie_chart)
```

-----------------------------------------------------------------------

# Libraries
library(tidyverse)
library(maps)
library(sf)
library(shiny)
library(plotly)

# Function to create the interactive map
createInteractiveMap <- function(selected_country) {
  # Import the data with coordinates
  world_map <- map_data("world")
  accesstonet <- read.csv("food-waste-by-country-2023.csv")
  map_data <- accesstonet %>% rename(country = Country)

  map <- left_join(world_map, map_data, by = "country")

  # Create the ggplot object
  gg <- ggplot(map, aes(x = long, y = lat, group = group, text = paste("Country: ", country, "<br>Annual Kg per Capita: ", FoodWaste2021kgcapitayear))) +
    geom_polygon(aes(fill = FoodWaste2021kgcapitayear), color = "white", size = 0.03) +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    theme_minimal()

  # Convert ggplot to plotly
  map_plotly <- ggplotly(gg, tooltip = "text")

  # Set initial zoom to the selected country
  if (!is.null(selected_country)) {
    map_plotly$x$layout$geo$fitbounds <- "locations"
    map_plotly$x$data[[1]]$locations <- selected_country
  }

  # Display the interactive map
  return(map_plotly)
}

# Shiny app
ui <- fluidPage(
  selectInput("selected_country", "Select Country", choices = ""),
  plotlyOutput("interactive_map")
)

server <- function(input, output, session) {
  observe({
    choices <- unique(read.csv("food-waste-by-country-2023.csv")$Country)
    updateSelectInput(session, "selected_country", choices = choices)
  })

  # Update the map based on the selected country
  output$interactive_map <- renderPlotly({
    selected_country_data <- filter(read.csv("food-waste-by-country-2023.csv"), country == input$selected_country)
    map_plot <- createInteractiveMap(selected_country_data$country)
    map_plot
  })
}

shinyApp(ui, server)


```

#intend to able to zoom into country when choosing the country,default singapore

### Dataset

```{r}
# Libraries
library (tidyverse)
library (plotly)
library (shiny)
library(dplyr)

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

energy_stat <- read.csv('waste_energy_stat.csv')


# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Filtering waste_03_22 dataset for specific years and waste types
wasteselected <- clean_waste_03_22 %>%
  filter(year %in% 2003:2022,
         waste_type %in% c("Plastics", "Ferrous Metals", "Non-Ferrous Metals", "Glass", "Food"))%>%
  arrange(waste_type)

  
### Calculating and adding recycling rate into DataFrame for analysis.

# Calculate recycling_rate and round to 2 decimal places
clean_waste_03_22$recycling_rate <- round(
  clean_waste_03_22$total_waste_recycled_tonne / clean_waste_03_22$total_waste_generated_tonne,2
)

### Calculating and adding wasting rate into DataFrame for analysis.

# Calculate wasting_rate and round to 2 decimal places
clean_waste_03_22$wasting_rate <- round(
  1-clean_waste_03_22$recycling_rate,2
)

# Display the cleaned data
head(clean_waste_03_22)
```


```{r}
# Libraries
library(tidyverse)
library(shiny)
library(plotly)

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Define UI for the Shiny app
ui <- fluidPage(
  titlePanel("Waste Visualization App"),
  sidebarLayout(
    sidebarPanel(
      selectInput("material", "Choose Material:", 
                  choices = unique(clean_waste_03_22$waste_type),
                  selected = "Plastics")
    ),
    mainPanel(
      plotlyOutput("waste_plot")
    )
  )
)

# Define server logic for the Shiny app
server <- function(input, output) {
  
  # Filter data based on selected material
  selected_data <- reactive({
    clean_waste_03_22 %>%
      filter(waste_type == input$material)
  })
  
  # Render the plot
  output$waste_plot <- renderPlotly({
    gg <- ggplot(selected_data(), aes(x = year)) +
      geom_line(aes(y = total_waste_generated_tonne, color = "Waste Generated"), size = 1) +
      geom_line(aes(y = total_waste_recycled_tonne, color = "Waste Recycled"), size = 1) +
      geom_line(aes(y = total_waste_not_recycled_tonne, color = "Waste Not Recycled"), size = 1) +
      labs(title = paste("Waste Over Years -", input$material),
           x = "Year", y = "Tonnes") +
      theme_minimal()
    
    ggplotly(gg, tooltip = "text")
  })
}

# Run the Shiny app
shinyApp(ui, server)
```
```{r}
# Libraries
library(tidyverse)
library(shiny)
library(plotly)

# Reading the CSV files
waste_03_22 <- read.csv('2003_2022_waste.csv')

# Renaming columns and mutating the data to convert values from kilo-tonnes to tonnes
clean_waste_03_22 <- waste_03_22 %>%
  rename(
    waste_type = "waste_type",
    total_waste_generated_tonne = "total_generate_1k_tonnes",
    total_waste_recycled_tonne = "total_recycled_1k_tonnes",
    total_waste_not_recycled_tonne = "total_not_recycled_1k_tonnes",
    year = "year"
  ) %>%
  mutate(
    total_waste_generated_tonne = total_waste_generated_tonne * 1000,
    total_waste_recycled_tonne = total_waste_recycled_tonne * 1000
  )

# Filtering waste_03_22 dataset for specific years and waste types
wasteselected <- clean_waste_03_22 %>%
  filter(year %in% 2003:2022,
         waste_type %in% c("Plastics", "Ferrous Metals", "Non-Ferrous Metals", "Glass", "Food")) %>%
  arrange(waste_type)

  
### Calculating and adding recycling rate into DataFrame for analysis.

# Calculate recycling_rate and round to 2 decimal places
clean_waste_03_22$recycling_rate <- round(
  clean_waste_03_22$total_waste_recycled_tonne / clean_waste_03_22$total_waste_generated_tonne, 2
)

### Calculating and adding wasting rate into DataFrame for analysis.

# Calculate wasting_rate and round to 2 decimal places
clean_waste_03_22$wasting_rate <- round(
  1 - clean_waste_03_22$recycling_rate, 2
)

# Define UI for the Shiny app
ui <- fluidPage(
  titlePanel("Recycling and Wasting Rates Visualization App"),
  sidebarLayout(
    sidebarPanel(
      selectInput("material", "Choose Material:",
                  choices = unique(clean_waste_03_22$waste_type),
                  selected = "Plastics")
    ),
    mainPanel(
      plotlyOutput("rate_plot")
    )
  )
)

# Define server logic for the Shiny app
server <- function(input, output) {
  
  # Filter data based on selected material
  selected_data <- reactive({
    clean_waste_03_22 %>%
      filter(waste_type == input$material)
  })
  
  # Render the plot
  output$rate_plot <- renderPlotly({
    gg <- ggplot(selected_data(), aes(x = year)) +
      geom_line(aes(y = recycling_rate, color = "Recycling Rate"), size = 1) +
      geom_line(aes(y = wasting_rate, color = "Wasting Rate"), size = 1) +
      labs(title = paste("Recycling and Wasting Rates Over Years -", input$material),
           x = "Year", y = "Rate") +
      theme_minimal()
    
    ggplotly(gg, tooltip = "text")
  })
}

# Run the Shiny app
shinyApp(ui, server)
```


### Chart B

```{r}
# Colors
custom_colors <- viridis::mako(n=30)

# Grouping by material and summarizing total waste and recycled
summary_df <- clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated))

# Adding relevant headings
colnames(summary_df) <- c("Material", "Total Waste Generated (tonnes)", "Total Waste Recycled (tonnes)")

# Displaying the summarized data
summary_df

# Adapted code for waste data
clean_waste_03_22 %>%
  group_by(waste_type) %>%
  summarise(
    total_waste_generated = sum(total_waste_generated_tonne),
    total_waste_recycled = sum(total_waste_recycled_tonne)
  ) %>%
  arrange(desc(total_waste_generated)) %>%
  head(15) 

```

### Chart C

```{r}
# Bar Plot for Waste Generated and Waste Recycled
fig <- plot_ly(
  x = clean_waste_03_22$year,
  y = clean_waste_03_22$total_waste_generated_tonne,
  type = "bar",
  name = "Waste Generated"
) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_recycled_tonne,
    type = "bar",
    name = "Waste Recycled"
  ) %>%
  add_trace(
    x = clean_waste_03_22$year,
    y = clean_waste_03_22$total_waste_not_recycled_tonne,
    type = "bar",
    name = "Waste Not Recycled"
  )

fig
```

### Chart D

```{r}
#use merge function: 
total_data <- merge(clean_waste_03_22, energy_stat, by.x = "waste_type", by.y = "material", all.x = TRUE)
total_data$energy_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$energy_saved)
total_data$crude_oil_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$crude_oil_saved)

# Display the first few rows
head(total_data)

# Box plot of total_waste_not_recycled_tonne
fig <- plot_ly(total_data, x = ~year, y = ~total_waste_recycled_tonne, type = "box")
fig <- fig %>% layout(title = "Box Plot of Total Waste Not Recycled Tonne", yaxis = list(title = "Total Waste Not Recycled Tonne"))

fig

# Box plot of total_energy_saved
fig2 <- plot_ly(total_data, x = ~year, y = ~energy_saved, type = "box")
fig2 <- fig2 %>% layout(title = "Box Plot of Total Energy Saved", yaxis = list(title = "Total Energy Saved"))

fig2
```

### Chart E

```{r}
#scatter plot graph
#use merge function: 
total_data <- merge(clean_waste_03_22, energy_stat, by.x = "waste_type", by.y = "Material", all.x = TRUE)
total_data$energy_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$energy_saved)
total_data$crude_oil_saved <- total_data$total_waste_not_recycled_tonne * as.numeric(energy_stat$crude_oil_saved)

fig3 <- plot_ly(
  data = total_data,
  x = ~year,
  y = ~energy_saved,
  size = ~total_waste_generate_tonne,
  color = ~material,
  sizes = c(10, 60),
  type = "scatter",
  mode = "markers"
)

fig3


# Energy saved per year
annual_energy_savings <- total_data %>%
  group_by(year) %>%
  summarize(energy_saved = sum(energy_saved))

# Convert total_energy_saved to GWh
annual_energy_savings$total_energy_saved <- round(annual_energy_savings$energy_saved / 1000000, 2)

# Display the result
tail(annual_energy_savings)
```

